Using reference models in feature selection problems \vspace{.1in}

### 2019-10-17 paper/ref_approach.tex
clean up sections 1 to 3

\begin{document}
\maketitle
\begin{abstract}
  Feature selection, or more generally, model reduction is an important aspect of statistical workflow aiming to provide insights from data. In this paper we discuss and demonstrate the benefits of using a reference model in feature selection. A reference model acts as a noise-filter on the target variable by modeling its data generating mechanism. Use of the reference model predictions in the model selection reduces the variability and improves stability leading to improved model selection performance. Assuming that a Bayesian reference model describes well the future data distribution, it can be shown that decision theoretically justified approach is to project the reference model to a reduced model leading to projection predictive variable selection approach. The reference model can be used also ad-hoc in combination with common feature selection methods.  In several numerical experiments, we demonstrate superior performance of projection predictive approach and show how use of a reference model translates into better and more stable feature selection independently of the specific selection procedure. 
% These benefits are also illustrated on a real-world example of predicting the amount of body fat for which we compare a reference model based selection method to a purely data driven approach.
\end{abstract}

@@

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward
regression \citep{venables2013modern,harrell2015regression}, the
observed empirical data distribution is utilised as a proxy of future
data usually in combination with cross-validation or information
criteria to provide estimates for out-of-sample predictive
performance.  In contrast, reference model based methods approximate
the future data generation mechanism using the predictive distribution
of a reference model, which can be, for example, a full-encompassing
model including all covariates.  The main assumption under any
reference model approach is that we operate in an
$\mathcal{M}$-\textit{complete} framework
\citep{book:bernardo_smith,vehtari2012survey}, that is, we have
constructed a model which reflects in the best way our beliefs about
the future data and which has passed model checking and criticism
\citep[see, e.g.][]{gelman2013bayesian}.  The reference model approach
has been used for long time in Bayesian statistics since the seminal
work of \citet{paper:reference_lindley}. For more historical
references, see papers by \citet{vehtari2012survey} and
\citet{paper:model_selection}, and for most recent methodological
development see \citet{paper:projpred}.

@@

For the selection via projpred, the Bayesian reference model includes
all the covariates using a regularised horseshoe prior on the
covariate coefficients \citep{paper:rhs}. Submodels are explored using
forward search (the results are not sensitive to whether forward or
backward search is used), and the predictive utility is the expected
log-predictive density (elpd) estimated using PSIS-LOO
\citep{paper:psis_loo}.  We select the submodel as the one with the
smallest size which has an elpd score similar to the reference model
when taking into account the uncertainty in estimating the predictive
model performance. See Appendix A for brief review of projection
predictive approach and papers by \citet{paper:model_selection} and
\citet{paper:projpred} for more details. projpred selection is made
using the \texttt{projpred} package.

@@

First two rows of Table \ref{tab:model_performances} report the
predictive performances, in terms of cross-validated root mean square
error (RMSE), of the full model and the selected models using either
the projection predictive approach or the stepwise backward
elimination.  There is no difference in predictive performance of the
selected models by different approaches even if there is difference in
the selected features.

We repeat the experiment with a modified data set by adding 84
unrelated noisy features. Last two rows of Table
\ref{tab:model_performances} report cross-validated RMSE, the size of
the selected model and the number of selected noisy features using
either the projection predictive approach or the stepwise backward
elimination.  The projection predictive approach has similar
predictive performance and the same submodel size as in the first
experiment, whereas the stepwise regression has worse predictive
performance and selects a much larger submodel including a large
number of irrelevant features.

@@

Both projpred and steplm compare a large number of models using either
forward or backward search, which can lead to selection induced
overfitting, but even with 100 covariates projpred is able to select a
submodel with similar performance as the full model. Although projpred
uses Bayesian inference and steplm maximum likelihood, it is not
here the decisive feature explaining the difference in the performance.  The
main difference between the approaches is whether a reference model is
used or not, and in the rest of the paper we look into more detail how
the benefits and limitations of the reference model approach.

\subsection{Benefits and cost of using a reference model}

@@

In the bodyfat example above we desired at the same time good
predictive performance and selection of a smaller number of
features. When the goal is to select a minimal subset of features
which have similar predictive performance as the full model we call it
\emph{minimal subset feature selection}. This minimal subset might
exclude features which have some predictive information about the
target, but given the minimal subset these features are nit able to
provide such additional information which would improve predictive
performance. The usual reason for this is that the relevant features
not in the minimal subset highly correlate with features already in
the minimal subset. We will later return to a problem of finding all
the features with some predictive power.

Using the data generating mechanism \eqref{eq:simulated_data}, we
simulate data sets of different size with a relatively large number of
features $p=100$ with with $k=20$ of them being predictive.
We compare the minimal subset feature selection perfomance of
projection predictive approach (with a reference model), a Bayesian
stepwise forward selection (without a reference model), and stepwise
forward selection with and without a reference model.
\begin{itemize}
\item projpred: The reference model is a Bayesian linear regression
  model (ie, the posterior predictive distributions are used) using
  the first five supervised principal components. The search heuristic
  is forward search, and the predictive performance is estimated via
  10-fold cross-validation.
\item steplm (data based): Stepwise selection using AIC as in the bodyfat example.
\item steplm using a reference model: The reference model is the same as for projpred, but only point predictions (posterior predictive means) $\hat{y}$ are used to replace the target variable $y$. Same AIC selection is used in data based steplm.
\item Bayesian stepwise selection (data based): At each step the model
  fitted is a Bayesian linear regression using regularised horseshoe
  prior and the feature excluded is the one with higher Bayesian
  $p$-value, defined as:
\begin{equation}
\text{Bayesian p-value}(\theta) := \text{min}\{P(\theta\leq0|D),P(\theta>0|D)\}
\end{equation}
The selection continues if the reduced model has an elpd score higher than the
current fitted model.
\item Bayesian stepwise selection (ref model based): same as the data based, but instead of using the observed data values to fit the models at each step, it uses the observed covariate and the target values predicted by a reference model 
\end{itemize}

In Figure \ref{fig:rmse_vs_fdr} it is shown, for different values of
$n$ and $\rho$, the predictive performance in terms of root mean
square error and the false discovery rate of the selected submodel,
averaged after 100 data simulations. projpred has superior FDR and the
smallest RMSE. Using a reference model improves steplm significantly,
but the minimal subset feature selection stays unreliable. Stepwise
Bayesian linear regression model without a reference model is better
than steplm without a reference model, but the minimal subset feature
selection is unreliable.

Figure \ref{fig:entropy} reports the entropy scores for the different
methods compared. Lower entropy corresponds to a more stable
selection. Highly correlated predictive features may happen to be
alternately selected, making the measure of the stability of the
selection a non-trivial task. Entropy suffers of such a problem, thus
such a measure should be seen as one of different possible
measurements of the actual stability. The use of a reference model
improves slightly stability of steplm in feature selection. The
projection predictive approach results to be far more stable than all
the other ones. This is likely due to as projpred is based on better
decision theoretical formulation which 1) takes into account the full
predictive distribution and not just point estimate and 2) projects
the reference model posterior to the submodel instead of using a
simple refit of submodels.

@@

% \hypertarget{comparison}{%
%   \section{A comparison in the normal means problem framework}\label{comparison}}
\section{Complete feature selection}

>>>>

\begin{document}
\maketitle
\begin{abstract}
  Feature selection, or more generally, model reduction is an important aspect of the statistical workflow aiming to provide insights from data. In this paper we discuss and demonstrate the benefits of using a reference model in feature selection. A reference model acts as a noise-filter on the target variable by modeling its data generating mechanism. As a result, using the reference model predictions in the model selection procedure reduces the variability and improves stability leading to improved model selection performance. Assuming that a Bayesian reference model describes the true distribution of future data well, the theoretically preferred usage of the reference model is to project its predictive distribution to a reduced model leading to projection predictive variable selection approach. Alternatively, reference models may also be used in an ad-hoc manner in combination with common feature selection methods.  In several numerical experiments, we demonstrate superior performance of projection predictive approach and show how use of a reference model translates into better and more stable feature selection independently of the specific selection procedure. 
% These benefits are also illustrated on a real-world example of predicting the amount of body fat for which we compare a reference model based selection method to a purely data driven approach.
\end{abstract}

@@

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward
regression \citep{venables2013modern,harrell2015regression}, the
observed empirical data distribution is utilised as a proxy of future
data usually in combination with cross-validation or information
criteria to provide estimates for out-of-sample predictive
performance.  In contrast, reference model based methods approximate
the future data generation mechanism using the predictive distribution
of a reference model, which can be, for example, a full-encompassing
model including all covariates.  The main assumption under any
reference model approach is that we operate in an
$\mathcal{M}$-\textit{complete} framework
\citep{book:bernardo_smith,vehtari2012survey}, that is, we have
constructed a model which reflects our beliefs about
the future data in the best possible way and which has passed model checking and criticism
\citep[see, e.g.][]{gelman2013bayesian}.  The reference model approach
has been used for a long time in Bayesian statistics since the seminal
work of \citet{paper:reference_lindley}. For more historical
references, see the papers by \citet{vehtari2012survey} and
\citet{paper:model_selection}, and for most recent methodological
development see \citet{paper:projpred}.

@@

For the selection via projpred, the Bayesian reference model includes
all the covariates using a regularised horseshoe prior on the
covariate coefficients \citep{paper:rhs}. Submodels are explored using
forward search (the results are not sensitive to whether forward or
backward search is used), and the predictive utility is the expected
log-predictive density (elpd) estimated using approximate leave-one-out cross-validation via Pareto-smoothed importance-sampling
\citep[PSIS-LOO-CV; ][]{paper:psis_loo}.  We select the smallest  submodel with an elpd score similar to the reference model
when taking into account the uncertainty in estimating the predictive
model performance. See Appendix A for brief review of projection
predictive approach and \citet{paper:model_selection} and
\citet{paper:projpred} for more details. The complete projpred procedure is implemented in the \texttt{projpred} R package.

@@

In the first two rows of Table \ref{tab:model_performances}, we report the
predictive performances, in terms of cross-validated root mean square
error (RMSE), of the full model and the selected models using either
the projection predictive approach or the stepwise backward
elimination.  There is no difference in predictive performance of the
selected models by different approaches even if there is difference in
the selected features.

We repeat the experiment with a modified data set by adding 84
unrelated noisy features. In the last two rows of Table
\ref{tab:model_performances}, we report the cross-validated RMSE, the size of
the selected model and the number of selected noisy features using
either the projection predictive approach or the stepwise backward
elimination.  The projection predictive approach has similar
predictive performance and the same submodel size as in the first
experiment, whereas the stepwise regression has worse predictive
performance and selects a much larger submodel including a large
number of irrelevant features.

@@

Both projpred and steplm compare a large number of models using either
forward or backward search, which can lead to selection induced
overfitting, but even with 100 covariates, projpred is able to select a
submodel with similar performance as the full model. In this example, the two
 compared methods also differ in other aspects than the usage of a reference 
 model, such as that projpred
uses Bayesian inference and steplm uses maximum likelihood estimation. 
However, as we show in the present paper, one of the primary difference 
with respect to quality of the variable selection is indeed whether or 
not a reference model approach is applied.

\subsection{Benefits and costs of using a reference model}

@@

In the bodyfat example above we desired at the same time good
predictive performance and selection of a smaller number of
features. When the goal is to select a minimal subset of features
which have similar predictive performance as the full model we call it
\emph{minimal subset feature selection}. This minimal subset might
exclude features which have some predictive information about the
target but given the minimal subset, these features are not able to
provide such additional information which would improve predictive
performance. The usual reason for this is that the relevant features
not in the minimal subset highly correlate with features already in
the minimal subset. We will later return to a problem of finding all
the features with some predictive power.

Using the data generating mechanism \eqref{eq:simulated_data}, we
simulate data sets of different size with a relatively large number of
features $p=100$ with with $k=20$ of them being predictive.
We compare the minimal subset feature selection perfomance of
projection predictive approach (with a reference model), a Bayesian
stepwise forward selection (without a reference model), and stepwise
forward selection with and without a reference model.
\begin{itemize}
\item projpred: The Projective prediction approach is used. 
  The reference model is a Bayesian linear regression
  model using the first five supervised principal components as predictors
  and the full posterior predictive distribution as the basis of the projection.   
  The search heuristic  is forward search and the predictive performance 
  is estimated via 10-fold cross-validation.

\item steplm (data based): Stepwise selection using AIC as in the bodyfat example.

\item steplm (reference model based): The reference model is the same as for
 projpred, but only point predictions (posterior predictive means) $\hat{y}$ 
 are used to replace the target variable $y$. The same AIC selection strategy 
 as in the data based steplm is used.

\item Bayesian stepwise selection (data based): At each step, the fitted model
 is a Bayesian linear regression using the regularised horseshoe
  prior and the feature excluded is the one with the highest Bayesian
  $p$-value defined as:
\begin{equation}
\text{Bayesian p-value}(\theta) := \text{min}\{P(\theta\leq0 \; | \; D),P(\theta>0 \; | \; D)\}
\end{equation}
The selection continues if the reduced model has an elpd score higher 
(i.e., better) than the current model.

\item Bayesian stepwise selection (reference model based): The reference model is the same as for projpred, but only point predictions (posterior predictive means) $\hat{y}$ are used to replace the target variable $y$. The same Bayesian
  $p$-value selection strategy as in the data based Bayesian stepwise selection is used.

%Same as the data based Bayesian stepwise selection, but instead of using the observed data values to fit the models at each step, it uses the target values predicted by the reference model (which is the same as in the projpred approach).
\end{itemize}

In Figure \ref{fig:rmse_vs_fdr}, the predictive performance is shown
for different values of $n$ and $\rho$ in terms of RMSE and the false discovery rate (FDR) of the selected submodel,
averaged after 100 data simulations. We see that projpred has superior FDR and the
smallest RMSE. Using a reference model improves steplm significantly,
but the minimal subset feature selection stays unreliable. Stepwise
Bayesian linear regression model without a reference model is better
than steplm without a reference model, but the minimal subset feature
selection is unreliable.

Figure \ref{fig:entropy} reports the entropy scores for the different
methods compared. Lower entropy corresponds to a more stable
selection. Highly correlated predictive features may happen to be
alternately selected, making the measure of the stability of the
selection a non-trivial task. Entropy suffers of such a problem, thus
such a measure should be seen as one of different possible
measurements of the actual stability. 
[TODO: please clarify the above sentences about entropy.] 
The use of a reference model
improves slightly stability of steplm in feature selection. The
projection predictive approach results to be far more stable than all
other approaches. This is likely due to projpred being based on better
decision theoretical formulation which 1) takes into account the full
predictive distribution and not just point estimate and 2) projects
the reference model posterior to the submodel instead of using a
simple refit of submodels.

@@

% \hypertarget{comparison}{%
%   \section{A comparison in the normal means problem framework}\label{comparison}}
\section{Complete feature selection}\label{comparison}
